The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
moved on to review section 4 (dynamics model)

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics},~\citep{lemaignan2014dynamics}) represents how the
level of anthropomorphic effects~\Ant[R,H,C,t] (\ie the observable
manifestations of anthropomorphism) evolves over a long-term human-robot
interaction.

@@

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects} $\antNorm[R,H,C,t] = \frac{\ant[R,H,C,t]}{\antMax}$:
because anthropomorphic effects are often qualitative and difficult to quantify
on an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted \AntMax, with
$t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). For the sake of readability,
we will use \Ant in the following sections to refer to the normalized level of
anthropomorphic effects. Note also that we discuss at
section~\ref{sec:measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

The actual maximum value of anthropomorphic effects depends on each unique
combination of human, robot and several other factors we introduce below, and
thus varies (as pictured on Figure~\ref{fig:dynamics}, we suggest besides that
$t_{max}$ actually corresponds to the novelty effect peak). The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
    users build an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user, the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphic effects.}

@@

We distinguish three main phases that describe the evolution of the
anthropomorphic effects in a long-term human-robot interaction. They are
depicted in different shades on Figure~\ref{fig:dynamics}.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} \SLA. \SLA may be null (no anthropomorphic
effects observed anymore), but it may also remain at a higher level.  This
third phase, as well as the \emph{stabilized level of anthropomorphism}, are
discussed in section~\ref{sec:stabilization} below.

@@

The \emph{initial capital of anthropomorphism} \ICA describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors, introduced in
section~\ref{sec:anthropomorphism}. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\citep{epley_when_2008}. Also the shape and design of the robot play a
role, like the context in which the interaction takes place. Our model of
anthropomorphism represents these factors as a compound metric \ICA. As such,
\ICA describes the level of anthropomorphic projections just prior the first
(real or imagined) contact to a robot. In this stage of pre-interaction, people
form initial expectations toward the robot and imagine how they will use it /
interact with it.

Following the discussion in section~\ref{sec:anthropomorphism}, we build \ICA on
three main factors that determine \emph{a priori} that a robot
will be anthropomorphized:

@@

    \item ~\emph{Human-centered factor} $S$: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\citep{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural background,
        professional background).

    \item ~\emph{Robot-centered factor} $R$: The robot's \textbf{design} and how it
        appears to the human user. Characteristics of the robot's form,
        behavior, and interaction modalities (anthropomorphic
        design)~\citep{fong_survey_2003}.

    \item ~\emph{Context-centered factor} $C$: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the
        robot is used / experienced (environmental
        context)~\citep{joosse_what_2013}.

@@

For \textbf{personality}, we suggest to apply \citet{epley_seeing_2007}
psychological \textit{3-factor theory} of anthropomorphism. For instance,
children generally tend to anthropomorphize objects more than adults. Also, a
person who lacks social connection is said to be more likely to
anthropomorphize. Both aspects would increase \ICA. This means that, one and the
same robot used by a different user, can lead to a different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design (\eg Nao) leads to a higher \ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the \ICA. Quantify the \emph{level} of anthropomorphic design in a robot, as
suggested for instance by \citet{fong_survey_2003}, can be difficult.
Nevertheless, the important aspect is how the robot appears to the user, \eg
human-like or machine-like.

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in \citet{joosse_what_2013}. The authors showed for instance that when the
same robot (Nao) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different personalities to the robot. In general,
a robot which is imagined in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities~\citep{baxter2013do}).

>>>>

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics},~\citep{lemaignan2014dynamics}) represents how the
level of `anthropomorphic effects'~\Ant[R,H,C,t] (\ie the observable
manifestations of anthropomorphism) evolves over a long-term human-robot
interaction.\footnote{We cannot make a precise statement about how long `long-term' interaction has to be to be considered as such. It seems to be practical to not speak of `several days' or `weeks' but to rather consider the duration and frequency of the actual interactions with the artifact. One could probably call this the `interaction history'.}

@@

In the \emph{Dynamics of Anthropomorphism} model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects} $\antNorm[R,H,C,t] = \frac{\ant[R,H,C,t]}{\antMax}$:
because anthropomorphic effects are often qualitative and difficult to quantify
on an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted \AntMax, with
$t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). For the sake of readability,
we will use \Ant in the following sections to refer to the normalized level of
anthropomorphic effects. Note also that we discuss in
Section~\ref{sec:measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

The actual maximum value of anthropomorphic effects depends on each unique
combination of a human, a robot, a situation and several other factors we introduce below. Thus, the maximum value varies according to the specific characteristics of these factors. Figure~\ref{fig:dynamics} pictures this maximum value at
$t_{max}$, which actually corresponds to what we call the novelty effect peak.\footnote{We assume that there is a peak in the human's experiences of the novelty effects of the unfamiliar system. Further, we suspect that this novelty peak comes along with the maximum value of anthropomorphic effects.} 
 The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.\fixme{Last sentence: Do you mean the shape of the curve? i would not agree that it remains the same ... but say that it shows a general tendency to decrease, however that there might be variations.}

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. The pre-interaction phase
    is characterized through an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user (and his/her expectations), the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphic effects.}

@@

We distinguish three main phases that describe the evolution of the
anthropomorphic effects in a long-term human-robot interaction. They are
depicted in different shades in Figure~\ref{fig:dynamics}.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} \SLA. \SLA may be null (no anthropomorphic
effects observed anymore), but it may also remain at a higher level.  This
third phase, as well as the \emph{stabilized level of anthropomorphism}, are
discussed in section~\ref{sec:stabilization}.

@@

The \emph{initial capital of anthropomorphism} \ICA describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors, introduced in
section~\ref{sec:multi-factors}, see also Figure \ref{fig:factors}. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\citep{epley_when_2008}. Also the shape and design of the robot play a
role, as well as the context in which the interaction takes place. Our model of
anthropomorphism represents these factors as a compound metric \ICA. As such,
\ICA describes the level of anthropomorphic projections just prior the first
(real or imagined) interaction with a robot. In this stage of pre-interaction, people
form \textbf{initial expectations} toward the robot and imagine how they will use it /
interact with it.

Taking up on our suggestion to have four main factors that impact on anthropomorphism, we propose that the initial capital of anthropomorphism can be build \emph{a priori}.
As the \ICA marks the very moment before the first real or imagined contact of the human with the robot, the \textit{interaction history} between the two players is empty; consequently the time factor $t$ is not taken into account when building the \ICA.
The specific \ICA value reflects how far a robot is likely to be anthropomorphized, based on the following three factors:

@@

    \item ~\emph{Human factor} $S$: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\citep{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural background,
        professional background).

    \item ~\emph{Robot factor} $R$: The robot's \textbf{design} and how it
        appears to the human user. Characteristics of the robot's form,
        behavior, and interaction modalities (anthropomorphic
        design)~\citep{fong_survey_2003}.

    \item ~\emph{Context factor} $C$: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the
        robot is used / experienced (environmental
        context)~\citep{joosse_what_2013}.

@@

For \textbf{personality}, we suggest to apply the
psychological \textit{3-factor theory} of anthropomorphism by \citet{epley_seeing_2007}. For instance,
children generally tend to anthropomorphize objects more than adults. Also, a
person who lacks social connection is said to be more likely to
anthropomorphize. Both aspects would increase \ICA. This means that, one and the
same robot used by a different user, can lead to a different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design up to a certain acceptable level (\eg Nao) leads to a higher \ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the \ICA. Quantifying the \emph{level} of anthropomorphic design in a robot, as
suggested for instance by \citet{fong_survey_2003}, can be difficult.
Nevertheless, the important aspect is how the robot appears to the user, \eg
human-like or machine-like, acceptable or not-acceptable.\footnote{It becomes clear that the \textit{robot factor} can not be fully separated from the \textit{human factor}. The reason for this is that the message that design tries to make, always lies in the eye of the beholder. No matter how much effort a designer spends in making the `message' of the artifact's design clear, there is always some scope and creative freedom in how the design is interpreted by the beholder.}

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined application context of the robot, or in which situation the interaction occurs, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in \citet{joosse_what_2013}. The authors showed for instance that when the
same robot (Nao) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different personalities to the robot. In general,
a robot which is imagined in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This was also found in an analysis of anthropomorphic language in online forum discussions about the AIBO, Roomba, and iPad \citep{fink_anthropomorphic_2012}. Regardless of the conversation topic, more anthropomorphic language was found in the AIBO forum. However, this was not only due to the zoomorphic (pet-like) shape of AIBO but also due to its  playful purpose. The idea that the purpose of a robot are crucial factors, also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities~\citep{baxter2013do}).
